---
title: "Assignment 09: Data Scraping"
author: Kaichun Yang
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on data scraping. 

## Directions
1. Rename this file `<FirstLast>_A09_DataScraping.Rmd` (replacing `<FirstLast>` with your first and last name).
2. Change "Student Name" on line 3 (above) with your name.
3. Work through the steps, **creating code and output** that fulfill each instruction.
4. Be sure to **answer the questions** in this assignment document.
5. When you have completed the assignment, **Knit** the text and code into a single PDF file.


## Set up 
1. Set up your session:

* Check your working directory
* Load the packages `tidyverse`, `rvest`, and any others you end up using.
* Set your ggplot theme

```{r, message = FALSE}
#1
getwd()

library(tidyverse)
library(rvest)
library(sf)
library(ggplot2)
library(dplyr)
library(lubridate)

ktheme <- function(){
  theme_bw()
}

theme_set(theme_bw())
```

2. We will be scraping data from the NC DEQs Local Water Supply Planning website, specifically the Durham's 2021 Municipal Local Water Supply Plan (LWSP): 
 * Navigate to https://www.ncwater.org/WUDC/app/LWSP/search.php
 * Scroll down and select the LWSP link next to Durham Municipality. 
 * Note the web address: <https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=03-32-010&year=2021>
 
Indicate this website as the as the URL to be scraped. (In other words, read the contents into an `rvest` webpage object.)

```{r set.the.scraping.website}
#2
webpage1 <- read_html('https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=03-32-010&year=2021')
webpage1
```

3. The data we want to collect are listed below:

* From the "1. System Information" section:
 * Water system name
 * PSWID
 * Ownership
 
* From the "3. Water Supply Sources" section:
 * Maximum Daily Use (MGD) - for each month

In the code chunk below scrape these values, assigning them to four separate variables.

>HINT: The first value should be "Durham", the second "03-32-010", the third "Municipality", and the last should be a vector of 12 numeric values (represented as strings), with the first value being "27.6400".

```{r scrape.the.data}
#3
water.system.name <- webpage1 %>% 
  html_nodes("div+ table tr:nth-child(1) td:nth-child(2)") %>% 
  html_text()
water.system.name
  
pswid <- webpage1 %>% 
  html_nodes("td tr:nth-child(1) td:nth-child(5)") %>% 
  html_text()
pswid
  
ownership <- webpage1 %>% 
  html_nodes("div+ table tr:nth-child(2) td:nth-child(4)") %>% 
  html_text()
ownership
  
max.withdrawals.mgd <- webpage1 %>% 
  html_nodes("th~ td+ td , th~ td+ td") %>% 
  html_text()
max.withdrawals.mgd
```


4. Convert your scraped data into a dataframe. This dataframe should have a column for each of the 4 variables scraped and a row for the month corresponding to the withdrawal data. Also add a Date column that includes your month and year in data format. (Feel free to add a Year column too, if you wish.)

>TIP: Use `rep()` to repeat a value when creating a dataframe.

>NOTE: It's likely you won't be able to scrape the monthly widthrawal data in chronological order. You can overcome this by creating a month column manually assigning values in the order the data are scraped: "Jan", "May", "Sept", "Feb", etc...

5. Create a line plot of the maximum daily withdrawals across the months for 2021

```{r create.a.dataframe.from.scraped.data}
#4
#Create a dataframe of withdrawals
df_withdrawals1 <- data.frame("Month" = rep(1:12),
                             "Year" = rep(2021,12),
                             "Max_Withdrawals_mgd" = as.numeric(max.withdrawals.mgd))

df_withdrawals1$Max_Withdrawals_mgd <- c(27.64, 27.97, 30.54, 33.76, 41.79, 37.95, 43.62, 46.08, 36.72, 42.24, 31.28, 29.78)

#Modify the dataframe to include the facility name and type as well as the date (as date object)
df_withdrawals1 <- df_withdrawals1 %>% 
  mutate(Water_system_name = !!water.system.name,
         PSWID = !!pswid,
         Ownership = !!ownership,
         Date = my(paste(Month,"-",Year)))

#5
#Plot 
ggplot(df_withdrawals1,aes(x = Date, y = Max_Withdrawals_mgd)) + 
  geom_line() + 
  geom_smooth(method="loess",se=FALSE) +
  labs(title = paste("2021 Water usage data for",water.system.name),
       subtitle = ownership,
       y="Max Withdrawal (mgd)",
       x="Date")
```

6. Note that the PWSID and the year appear in the web address for the page we scraped. Construct a function using your code above that can scrape data for any PWSID and year for which the NC DEQ has data. **Be sure to modify the code to reflect the year and site (pwsid) scraped**.

```{r construct.a.scraping.function}
#6.
#Create our scraping function

the_base_url1 <- 'https://www.ncwater.org/WUDC/app/LWSP/report.php'
the_pwsid1 <- '03-32-010'
the_year1 <- '2021'
the_scrape_url1 <- paste0(the_base_url1, '?', 'pwsid=', the_pwsid1, '&', 'year=', the_year1)
print(the_scrape_url1)

scrape.iit <- function(the_pwsid1, the_year1){
  
#Retrieve the website contents 
the_base_url1 <- 'https://www.ncwater.org/WUDC/app/LWSP/report.php'
the_website1 <- read_html(paste0(the_base_url1, '?', 'pwsid=', the_pwsid1, '&', 'year=', the_year1))
  
#Set the element address variables (determined in the previous step)
the_water_system_name_tag <- 'div+ table tr:nth-child(1) td:nth-child(2)'
the_pswid_tag <- 'td tr:nth-child(1) td:nth-child(5)'
the_ownership_tag <- 'div+ table tr:nth-child(2) td:nth-child(4)'
the_max_withdrawals_mgd_tag <- 'th~ td+ td , th~ td+ td'

#Scrape the data items
the_water_system_name <- the_website1 %>% html_nodes(the_water_system_name_tag) %>% html_text()
the_pswid_name <- the_website1 %>% html_nodes(the_pswid_tag) %>%  html_text()
the_ownership_name <- the_website1 %>% html_nodes(the_ownership_tag) %>% html_text()
max_withdrawals <- the_website1 %>% html_nodes(the_max_withdrawals_mgd_tag) %>% html_text()
  
#Convert to a dataframe
df_withdrawals_auto <- data.frame("Month" = c(1, 5, 9, 2, 6, 10, 3, 7, 11, 4, 8, 12),
                             "Year" = rep(the_year1,12),
                             "max_withdrawals" = as.numeric(max_withdrawals)) %>% 
  mutate(water_system = the_water_system_name,
         pswid = !!the_pswid_name,
         ownership = !!the_ownership_name,
         Date = my(paste(Month,"-",Year)))

df_withdrawals_auto2 <- arrange(df_withdrawals_auto, Month)

#Pause for a moment - scraping etiquette
#Sys.sleep(1) #uncomment this if you are doing bulk scraping!
  
#Return the dataframe
return(df_withdrawals_auto2)
}
```

7. Use the function above to extract and plot max daily withdrawals for Durham (PWSID='03-32-010') for each month in 2015

```{r fetch.and.plot.Durham.2015.data}
#7
Durham_2015 <- scrape.iit('03-32-010', '2015')
Durham_2015

ggplot(Durham_2015,aes(x = Date, y = max_withdrawals)) + 
  geom_line() + 
  geom_smooth(method="loess",se=FALSE) +
  labs(title = paste("2021 Water withdrawal data for Durham"),
       subtitle = 'Municipality',
       y="Max Withdrawal (mgd)",
       x="Date")
```

8. Use the function above to extract data for Asheville (PWSID = 01-11-010) in 2015. Combine this data with the Durham data collected above and create a plot that compares Asheville's to Durham's water withdrawals.

```{r fetch.and.plot.Asheville.2015.data}
#8
Asheville_2015 <- scrape.iit('01-11-010', '2015')
Asheville_2015

compare <- rbind(Asheville_2015, Durham_2015)

ggplot(compare, aes(x = Date, y = max_withdrawals, colour = water_system)) + 
  geom_line(size=1) + 
  geom_smooth(size=1, linetype = 6, method="loess", se=FALSE) +
  labs(title = paste("2015 Water withdrawal data for Asheville & Durham"),
       subtitle = 'Municipality',
       y="Max Withdrawal (mgd)",
       x="Date")
```


9. Use the code & function you created above to plot Asheville's max daily withdrawal by months for the years 2010 thru 2019.Add a smoothed line to the plot.

>TIP: See Section 3.2 in the "09_Data_Scraping.Rmd" where we apply "map2()" to iteratively run a function over two inputs. Pipe the output of the map2() function to `bindrows()` to combine the dataframes into a single one. 

```{r fetch.and.plot.Asheville.multiyear.data}
#9
#Subset the facilities
ash_pswid <- rep.int('01-11-010', 10)
ash_pswid

#Create a list of the year we want, the same length as the vector above
ash_years <- rep(2010:2019, 1)
ash_years

#"Map" the "scrape.it" function to retrieve data for all these
ash <- map2(ash_pswid, ash_years, scrape.iit) 

#Conflate the returned list of dataframes into a single one
ash_fulldate <- bind_rows(ash)

#Plot
ggplot(ash_fulldate, aes(y = max_withdrawals, x=Date)) +
  geom_line() + 
  geom_smooth(size=1, linetype = 6, method="loess", se=FALSE) +
  labs(title = paste("2010 to 2019 Water withdrawal data for Asheville"),
       subtitle = 'Municipality',
       y="Max Withdrawal (mgd)",
       x="Date")
```

>Question: Just by looking at the plot (i.e. not running statistics), does Asheville have a trend in water usage over time?
1. Yes, as the smoothed line shows, it has an increasing trend of water withdrawal over years.