---
title: "Assignment 2: Coding Basics"
author: "Kaichun Yang"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics.

## Directions

1.  Rename this file `<FirstLast>_A02_CodingBasics.Rmd` (replacing `<FirstLast>` with your first and last name).
2.  Change "Student Name" on line 3 (above) with your name.
3.  Work through the steps, **creating code and output** that fulfill each instruction.
4.  Be sure to **answer the questions** in this assignment document.
5.  When you have completed the assignment, **Knit** the text and code into a single PDF file.
6.  After Knitting, submit the completed exercise (PDF file) to Sakai.

## Basics Day 1

1.  Generate a sequence of numbers from one to 100, increasing by fours. Assign this sequence a name.

2.  Compute the mean and median of this sequence.

3.  Ask R to determine whether the mean is greater than the median.

4.  Insert comments in your code to describe what you are doing.

```{r}
#1
### Use seq() to generate a sequence of numbers and assign the sequence as kc1
kc1 <- seq(1,100,4)
print(kc1)


#2
### Calculate mean and median of the sequence kc1 and assign them as kc2 and kc3
kc2 <- mean(kc1)
kc3 <- median(kc1)
print(kc2)
print(kc3)

#3 & #4
### Use if/else if/else to compare kc2, the mean of the sequence kc1 and kc3, 
### the median of the sequence kc1
### The function checks if the kc2 is bigger, if the kc3 is bigger, if kc2 and 
### kc3 are equal in order and when one of them is checked, the solution will be printed.
if (kc2>kc3) {
print('The mean is bigger')
} else if (kc2<kc3){
  print('The median is bigger')
} else {
  print('The mean and median are equal')
}


```

## Basics Day 2

5.  Create a series of vectors, each with four components, consisting of (a) names of students, (b) test scores out of a total 100 points, and (c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.

6.  Label each vector with a comment on what type of vector it is.

7.  Combine each of the vectors into a data frame. Assign the data frame an informative name.

8.  Label the columns of your data frame with informative titles.

```{r}
#5 & #6
### v1's type is character
### v2's type is double
### v3's type is character
v1 <- c('Alex', 'Brandon', 'Clara', 'David')
typeof(v1)
v2 <- c(33, 55, 77, 99)
typeof(v2)
v3 <- c('Flase', 'True', 'True', 'True')
typeof(v3)

#7 & #8
### Combine v1, v2, v3 together and assign titles
grades <- data.frame('Name' = v1, 'Scores' = v2, 'Pass' = v3)
class(grades)
print(grades)
```


9.  QUESTION: How is this data frame different from a matrix?

> Answer: Dataframe is able to contain different types of data, matrix has similar data type in its array. In other word, dataframe is a heterogeneous while matrix is homogeneous. Data frame is a generalized form of matrix and can have column and row names.


10. Create a function with an if/else statement. Your function should take a **vector** of test scores and print (not return) whether a given test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement. 

```{r}
#10
### The function 'grader' determine if the students pass or not, the function 'result'
###conduct grader and then print its result
result <- function(x){
grader <- function(x){
  ifelse(x < 50, 'FAIL', 'PASS')
}
print(grader(x))
}
```

11. Apply your function to the vector with test scores that you created in number 5.

```{r}
#11
### Use the created function for v2, the vector contains scores
result(v2)
```

12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked? Why?

> Answer: 'ifelse' works since it can process vectors. Other conditonal statements including if, else, if/else, and one line if...else are not vector operations but dealing with single value.
