---
title: "Assignment 5: Data Visualization"
author: "Student Name"
output: pdf_document
geometry: margin=2.54cm
fig_width: 6 
fig_height: 4
---

```{r setup, include=FALSE}
#This code chunk will tidy your knit PDF files, wrapping long code lines
#For it to work, the "formatR" package needs to be installed

#install.packages('formatR')
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE, echo = TRUE, warning = FALSE, message = FALSE)
```

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on Data Visualization 

## Directions
1. Rename this file `<FirstLast>_A02_CodingBasics.Rmd` (replacing `<FirstLast>` with your first and last name).
2. Change "Student Name" on line 3 (above) with your name.
3. Work through the steps, **creating code and output** that fulfill each instruction.
4. Be sure to **answer the questions** in this assignment document.
5. When you have completed the assignment, **Knit** the text and code into a single PDF file.

The completed exercise is due on Friday, Oct 14th @ 5:00pm.



## Set up your session 
1. Set up your session. Verify your working directory and load the tidyverse, 
lubridate, & cowplot packages. Upload the NTL-LTER processed data files for  
nutrients and chemistry/physics for Peter and Paul Lakes (use the tidy 
[`NTL-LTER_Lake_Chemistry_Nutrients_PeterPaul_Processed.csv`] version) and 
the processed data file for the Niwot Ridge litter dataset (use the
[`NEON_NIWO_Litter_mass_trap_Processed.csv`] version). 

2. Make sure R is reading dates as date format; if not change the format to date.

```{r}
#1 
library(tidyverse)
library(lubridate)
library(cowplot)
ntl = read.csv(file = 'E:/EDA-Fall2022/Data/Processed/NTL-LTER_Lake_Chemistry_Nutrients_PeterPaul_Processed.csv')
neon = read.csv(file = 'E:/EDA-Fall2022/Data/Processed/NEON_NIWO_Litter_mass_trap_Processed.csv')

#2 
neon_1 = neon
neon_1$collectDate <- as.Date(neon_1$collectDate)
```
## Define your theme
3. Build a theme and set it as your default theme.

```{r}
#3
kc_theme <- function(){
  theme_classic(base_size = 14) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top")
}
```

## Create graphs
For numbers 4-7, create ggplot graphs and adjust aesthetics to follow best 
practices for data visualization. Ensure your theme, color palettes, axes, 
and additional aesthetics are edited accordingly.

4. [NTL-LTER] Plot total phosphorus (`tp_ug`) by phosphate (`po4`), with 
separate aesthetics for Peter and Paul lakes. Add a line of best fit and 
color it black. Adjust your axes to hide extreme values (hint: change the 
limits using `xlim()` and/or `ylim()`).

```{r}
#4
class(ntl)
p0 <- ggplot(ntl, aes(tp_ug, po4), warning = FALSE)
p0 + geom_point(aes(colour = lakename)) + ylim(0,50)

```


5. [NTL-LTER] Make three separate boxplots of (a) temperature, (b) TP, and 
(c) TN, with month as the x axis and lake as a color aesthetic. Then, create 
a cowplot that combines the three graphs. Make sure that only one legend is 
present and that graph axes are aligned. 

Tip: R has a build in variable called `month.abb` that returns a list of months;
see <https://r-lang.com/month-abb-in-r-with-example>

```{r p7, fig.width = 12, fig.asp = .5}
#5

# p <- ggplot(ntl, lakename = 'group')
# p + geom_boxplot(aes(month, temperature_C), color=factor(group))

ntl$month <- as.factor(ntl$month)

p1 <- ggplot(ntl, aes(month, temperature_C)) + 
  geom_boxplot(aes(colour = lakename)) + 
  theme(legend.position = 'none')

p2 <- ggplot(ntl, aes(month, tp_ug)) + 
  geom_boxplot(aes(colour = lakename)) + 
  theme(legend.position = 'none')

p3 <- ggplot(ntl, aes(month, tn_ug)) + 
  geom_boxplot(aes(colour = lakename)) + 
  theme(legend.position = 'none')

p5 <- ggplot(ntl, aes(month, tn_ug)) + 
  geom_boxplot(aes(colour = lakename))

p4 <- plot_grid(p1, p2, p3, labels = c('A', 'B', 'C'), nrow = 3, align = 'hv') 

legend <- get_legend(p5 + theme(legend.box.margin = margin(0, 0, 0, 2)))

p7 <- plot_grid(p4, legend, rel_widths = c(3,1))

plot(p7)
```

Question: What do you observe about the variables of interest over seasons 
and between lakes? 

> Answer: Temperature doesn't change too much, Peter Lake always has higher tp_ug and tn_ng than Paul Lake 

6. [Niwot Ridge] Plot a subset of the litter dataset by displaying only the 
"Needles" functional group. Plot the dry mass of needle litter by date and 
separate by NLCD class with a color aesthetic. (no need to adjust the name 
of each land use)

7. [Niwot Ridge] Now, plot the same plot but with NLCD classes separated into 
three facets rather than separated by color. 

```{r p6, fig.width = 12, fig.asp = .5}
#6
neon_ndls <- filter(neon_1, functionalGroup == 'Needles')

neon_ndls$collectDate <- as.factor(neon_ndls$collectDate)

p6 <- ggplot(neon_ndls, aes(collectDate, dryMass)) + 
  geom_boxplot(aes(colour = nlcdClass)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

plot(p6)

#7
FacetPlot1 = ggplot(neon_ndls, aes(collectDate, dryMass)) + geom_boxplot() + facet_grid(~nlcdClass) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

plot(FacetPlot1)

```
Question: Which of these plots (6 vs. 7) do you think is more effective, and why? 

> Answer: 7 is more effective to see the trend of each dataset and 6 is more effectice to compare the valuse in each date. 